---
title: "Project title"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
CollegeBasketballPlayers2009_2021 <- read_csv("CollegeBasketballPlayers2009-2021.csv")

top_spotify_songs <- read_csv("songs_normalize 3.csv")

olympic_data <- read_csv("athlete_events.csv")
```

# Data 1

## Introduction and data

-   Identify the source of the data.

    The College Basketball Players 2009-2021 dataset is from Kaggle (https://www.kaggle.com/datasets/adityak2003/college-basketball-players-20092021)

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

    The data was collected by Bart Torvik (barttorvik.com)

-   Write a brief description of the observations.

    The observations give 65 variables related to college basketball for college basketball players from 2009-2021. Each of the 61,000 observations is one season by one player.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

How do various college basketball playing statistics influence players' NBA draft pick? Which statistic (efficiency rating, total points, shooting percentage, etc) is the best predictor of being a #1 or top 10 pick in the NBA draft?

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

The research topic is performance statistics of players in college basketball and the potential association with draft picks.

-   Identify the types of variables in your research question. Categorical? Quantitative?

The majority of the variables are quantitative such as average or total points, steals, blocks, etc but there are also some categorical variables, such as year in school, team, and conference.

## Literature

-   Find one published credible article on the topic you are interested in researching.

https://www.playingnumbers.com/2020/03/using-ncaa-stats-to-predict-nba-draft-order/

-   Provide a one paragraph summary about the article.

This article discussed two potential models which were created in attempts to predict draft order based on college statistics. One of the models included data for the players from only that season before the draft in question, while the other model included data of every season the player played in college before entering the draft. One strong predictor these models found was age, with freshman predicted and actually drafted higher on average and players drafted lower each subsequent year. However, one problem the article mentions in using college statistics to predict NBA draft picks is that because many prospects are drafted based on potential, there were often freshman with worse statistics that were drafted higher because NBA teams may have believed they had underdeveloped talent, as opposed to players with better statistics who the NBA teams may have viewed as having already reached their peak ability.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

Our research question is different than the article because it includes many more extensive variables than mentioned in the article, such as breaking down shots by dunks, rim shots, and midrange shots, along with categorizations such as conference which may influence players' success, so we will be able to look for more potential specific trends based on the numerous variables available.

## Glimpse of data

```{r}
#| label: load-data-1

glimpse(CollegeBasketballPlayers2009_2021)

```

# Data 2

## Introduction and data

-   Identify the source of the data.

    This is the Top Spotify hits from 2000-2019, top_spotify_songs, from Kaggle. https://www.kaggle.com/datasets/paradisejoy/top-hits-spotify-from-20002019

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

    The data was collected by Mark Koverha on Kaggle.

-   Write a brief description of the observations.

    This data set has 2000 observations and 16 columns of unique data sets (non-identifiers).

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

    1.  Do the most popular songs in any given year have the highest danceability ratings?

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

    This research topic is on the correlation between popular songs being the most danceable, and any trends in this over time. For this, our hypothesis is that the higher the danceability rating is, the more popular the song is going to be.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-2

glimpse(top_spotify_songs)
```

# Data 3

## Introduction and data

-   Identify the source of the data.

https://www.kaggle.com/datasets/heesoo37/120-years-of-olympic-history-athletes-and-results

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

Data was scraped and wrangled using r off of www.sports-reference.com by Randi H. Griffin.

-   Write a brief description of the observations.

-   

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-3

glimpse(olympic_data)
```
