---
title: "Project title"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
CollegeBasketballPlayers2009_2021 <- read_csv("CollegeBasketballPlayers2009-2021.csv")

top_spotify_songs <- read_csv("songs_normalize 3.csv")
```

# Data 1

## Introduction and data

-   Identify the source of the data.

    The College Basketball Players 2009-2021 dataset is from Kaggle (https://www.kaggle.com/datasets/adityak2003/college-basketball-players-20092021)

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

    The data was collected by Bart Torvik (barttorvik.com)

-   Write a brief description of the observations.

    The observations give 65 variables, from team to games played to shooting percentages and other statistics for 61,000 college basketball players from 2009-2021.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

How do various college basketball playing statistics influence players' NBA draft pick? Which statistic (efficiency rating, total points, shooting percentage, etc) is the best predictor of being a #1 or top 10 pick in the NBA draft?

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

https://www.playingnumbers.com/2020/03/using-ncaa-stats-to-predict-nba-draft-order/

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-1

glimpse(CollegeBasketballPlayers2009_2021)
```

# Data 2

## Introduction and data

-   Identify the source of the data.

    This is the Top Spotify hits from 2000-2019, top_spotify_songs, from Kaggle. https://www.kaggle.com/datasets/paradisejoy/top-hits-spotify-from-20002019

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

    The data was collected by Mark Koverha on Kaggle.

-   Write a brief description of the observations.

    This data set has 2000 observations and 16 columns of unique data sets (non-identifiers).

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-2

glimpse(top_spotify_songs)
```

# Data 3

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-3

# add code here
```
