---
title: "Project title"
subtitle: "Report"
format: html
editor: visual
execute:
  echo: true
---

# INTRODUCTION

## describe your topic (citing sources)

Our topic is the age, height, and weight of Olympic athletes in the sports of gymnastics and swimming. (idk what else to say)

## provide a concise, clear statement of your research question and hypotheses

How do factors like age, height, and weight correlate with medal success rate across gymnastics and swimming? We hypothesize that the ideal height, weight, and age to win a medal will all be greater in swimming than in gymnastics.

## consider any potential ethical issues that you believe should be addressed (if any) regarding how your data were collected

## identify the source of the data, when and how it was collected

This data is from www.sports-reference.com and was researched by a group of Olympic history enthusiasts. It was then compiled into the dataset format using R by Randi H. Griffin and uploaded to the Kaggle website in 2018.

## the cases

Each case in this dataset corresponds to an individual athlete competing in an individual Olympic event between the years of 1896 and 2016.

## a general description of relevant variables

For our research question, the relevant variables include the numerical variables age (in years), height (in centimeters), weight (in kilograms), and sport (character strings). For our analysis, we will filter the dataset to only include gymnastics and swimming, the sports we are interested in analyzing.

For this research topic, we would have to take a look at which ages, heights, and weights have received medals. I would hypothesize that for each separate sport, there is a different normal distribution for ages, heights, and weights that have medal success rates, with an ideal age, height, and weight for each individual sport.

# METHODOLOGY

visualizations and summary statistics relevant to your research question

justify the choice of statistical method(s) used to answer your research question

```{r}
#| label: packages
#| message: false
#| warning: false

library(tidyverse)
library(tidymodels)
```

```{r}
#| label: data
#| message: false
#| warning: false

olympic_data <- read_csv("athlete_events.csv")
```

```{r}
#| label: filter

olympic_data |>
  filter(Sport == "Swimming" | Sport == "Gymnastics")
```

# RESULTS

1.  select only the parts of the data set we need (filter by sports)
2.  mutate to get the medal values numeric
3.  visualize the height and weight and whether or not they received a medal
4.  then separate by gender and see if it changes

showcase how you arrived at answers to your research question using the techniques we have learned in class

provide only the main results from your analysis

demonstrate that you are proficient at asking meaningful questions and answering them using data

that you are skilled in interpreting and presenting results

and that you can accomplish these tasks using R
